---
title: "The Sky is Falling"
output: github_document
---

### The impact of title sentiment on the success of scientific YouTube content

```{r, message=FALSE, warning=FALSE}
library(tidyverse)
library(RPostgreSQL)
library(dotenv)
library(ggthemes)
library(ggthemr)
library(plotly)
library(formattable)

#securely loading credentials
load_dot_env(file = ".env")

#building a trycatch function for database connection
db_connect <- function(drv, host, port, dbname, user, password){
  tryCatch(
    dbConnect(drv=drv, host=host, port=port, dbname = dbname, user=user, password=password),
    error = function(e){
      print("Could not connect!") 
      print(e)
      },
    finally = {
      print("Connected!")
      }
  )
}

options(scipen = 999)
#connecting to the database
con <- db_connect(PostgreSQL(), host = Sys.getenv("ENDPOINT"),  port = Sys.getenv("PORT") , dbname = Sys.getenv("DB_NAME"), user = Sys.getenv("USERNAME"), password = Sys.getenv("PASSWORD"))
```

This is a followup on the NLP sentiment analysis on the titles for Kurzgesagt YouTube videos. In order to see how successful that sentiment analysis was, I will classify the titles manually and check to see how frequently the model agreed with my classifications. This data is clean, but I will add my subjective classifications as a new variable and I will ensure the upload date is classified as "date" data. 

```{r}
video_data <- collect(tbl(con, "video_data")) #collecting data from database to my system

#my sentiment ratings for the video titles
my_ratings <- c(-1, 1, 0, -1, 1, 0, -1, -1, 0, 0, 1, 0, 0, 1, 1, 0, 0, 0, 1, -1,-1, -1, -1, 0, -1, -1, -1, 1, -1, 0, -1, 0, 1, -1, 0, 1, 1, 0, 0, 0, -1, 0, 1, 0, 1, 0, -1, 0, -1, 1, 1, 1, 0, -1, -1, 0, 1, -1, -1, -1, -1, 0, -1, -1, -1, 1, 0, -1, 1, 0, 1, 1, -1, 0, 1, -1, 1, 0, -1, -1, 0, -1, -1, 1, 0, 0, -1, -1, -1, 1, 1, 1, 1, 0, -1, 0, -1, -1, -1, 1, 0, -1, 0, 0, 0, -1, -1, 1, -1, 1, -1, 0, 1, -1, 1, 1, 1, -1, 1, -1, -1, -1, 1, 0, 0, -1, 1, 0, -1, -1, 1, 0, 1, -1, 0, 0, -1, 0, 1, 0, -1, 0, 0, 0,-1, -1, 0, 0, 0, 0, 0, 0, 1, -1, 0, 0, 0, -1, 1, 1) %>% as.character #making character for easy replacement

my_ratings <- my_ratings %>%
  replace( my_ratings == "-1", "NEGATIVE") %>%
  replace(my_ratings == "0", "NEUTRAL") %>%
  replace(my_ratings == "1", "POSITIVE") 

video_data <- cbind(video_data, my_ratings)
video_data <- video_data %>%
  mutate(upload_date = as.Date(upload_date))
```

To check if the sentiment analysis model agrees with my assessment, I will first check the proportion of time that my title ratings matched the sentiment ratings of the model, and then I will filter out all "NEUTRAL" titles and repeat the process. 

```{r}
test1 <- video_data$title_sentiment == video_data$my_ratings
mean(test1)
#the model agreed with my rankings only 55% of the time

test2 <- video_data %>%
  filter(title_sentiment != "NEUTRAL" & my_ratings != "NEUTRAL") %>%
  mutate(test = ifelse(title_sentiment == my_ratings, 1, 0))
mean(test2$test)
#but if I remove neutral titles, it agrees with my rankings 85% of the time
```
When comparing directly, the model only agreed with my rankings 55% of the time, but when removing all "NEUTRAL" titles, the model agreed with my rankings 85% of the time (which is pretty good!). The issue is that the sentiment analysis model is really only designed to classify things as "POSITIVE" or "NEGATIVE". When I added the neutral class in the sentiment analysis, it was based on when the model had low certainty of either positive or negative, but it still wasn't enough to really capture neutral as a category. A way to improve this would be to train our own sentiment analysis model on a much larger set of video titles, which may be a longer-term project.  

```{r model-comparison-barplot}
ggthemr(palette = "fresh" ,layout = "pastel")
video_data %>%
  select(title_sentiment, my_ratings)%>%
  pivot_longer(cols = c("title_sentiment", "my_ratings"))%>%
  mutate( `Sentiment Source` = ifelse(name == "title_sentiment", "Model", "Human")) %>%
  ggplot(aes(x = value, fill = value))+
  geom_bar()+
  labs(x = 'Sentiment', y = "Number of Titles")+
  theme(strip.text.x = element_text(size = 14, face = "bold"), 
        strip.background = element_rect(fill = "grey99"), axis.text.x = element_text(face = "bold"), 
        axis.title = element_text(face = "bold"), legend.position = "none")+
  facet_wrap(~`Sentiment Source`)
```

As seen from the above visualization, the main difference between the model classifications and human classifications is that the model allocated a large portion of titles which I ranked as "NEUTRAL" to either the "POSITIVE" or "NEGATIVE" categories. 

A benefit of the sentiment analysis method is that it is not subject to personal bias - it has ranked certain words and collections of words as being associated with negative statements and some as being associated with positive statements. This is consistent even if it sometimes misses nuance. My own assessment of title sentiment is based on my subjective interpretation, which may change between evaluations. Both methods have pros and cons.

Now, doing some exploratory data analysis, I'd like to look at which titles are associated with the highest like, view, and comment counts. 

```{r, warning = F}
#Top 10 titles for views
video_data %>%  
  arrange(desc(view_count)) %>%
  rename(Title = vid_title, Views = view_count, `Sentiment (Modeled)` = title_sentiment) %>%
  select(Title,Views, `Sentiment (Modeled)`) %>%
  filter(Views >= Views[10]) %>%
  formattable()

#Top 10 titles for comments
video_data %>%  
  arrange(desc(comment_count)) %>%
  rename(Title = vid_title, Comments = comment_count, `Sentiment (Modeled)` = title_sentiment) %>%
  select(Title,Comments,`Sentiment (Modeled)` ) %>%
  filter(Comments >= Comments[10]) %>%
  formattable()

#Top 10 titles for likes
video_data %>%  
  arrange(desc(like_count)) %>%
  rename(Title = vid_title, Likes = like_count, `Sentiment (Modeled)` = title_sentiment) %>%
  select(Title,Likes, `Sentiment (Modeled)`) %>%
  filter(Likes >= Likes[10]) %>%
  formattable()
```

The video with the most views (by almost double the next highest) is an explanatory video which is discussing COVID and methods of mitigation. Interestingly, the video with the most likes and comments was not the video that had the most views (it had the 9th most). This video is called "The Egg - A Short Story", and is a non-scientific, hopeful and philosophical video that is not typical for this channel. This suggests that occasional deviations from your channel's main focus can still drive very high levels of positive engagement.

I'm curious if there is a difference in mean comments, likes, and views for the different titles. So let's investigate that:

```{r, warning = FALSE}
video_data %>%
  select(comment_count,like_count,view_count, my_ratings) %>%
  group_by(my_ratings) %>%
  rename(Sentiment = my_ratings) %>%
  summarize(`Mean Comments` = mean(comment_count),`Mean Likes` = mean(like_count), `Mean Views` = mean(view_count))%>%
  formattable
```
It appears that there is on average more engagement in the form of comments and likes for videos with either positive or negative titles, as opposed to neutral titles. Notably, the mean views were much higher (~19%) for videos with negative titles. So while engagement is similar between positive and negative titles, this could suggest that negative titles draw in more views overall.


In order to make sure this difference in mean views is statistically significant, a t-test between the positive and negative titles would likely be sufficient. A slightly more correct way to check though would be to perform an ANOVA along with a post-hoc analysis. 

```{r}
#ANOVA with Tukey post-hoc
views_anova <- aov(view_count~my_ratings, data = video_data)
TukeyHSD(views_anova)

#T test approach, investigating the difference in views between negative and positive title sentiment
ttest_data <- video_data%>%
  select(view_count, my_ratings)%>%
  filter(my_ratings != "NEUTRAL")
t.test(ttest_data$view_count~ttest_data$my_ratings)
```

Neither the anova or t test suggests a statistically significant difference in views between the two groups at an alpha of .05, however the t test p-value of .09 is low enough that this still warrants further investigation. A way to expand upon this would be to pull more data from additional science channels to see if this relationship becomes more apparent, which I will do in the next version of this project.


Let's visualize the differences between different title sentiments.

```{r violin-sentiment-plots, fig.width = 12}
ggthemr(palette = "pale", layout = "scientific")
#box plots of human data for likes, comments and views


video_data %>%
  select(view_count, like_count, comment_count, my_ratings) %>%
  rename(Views = view_count, Likes = like_count, Comments = comment_count)%>%
  pivot_longer(cols = c("Views", "Likes", "Comments")) %>%
  ggplot(aes(x = my_ratings, y = value))+
  geom_violin(aes(fill = my_ratings))+
  geom_boxplot(width = .2, fill = "grey95", color = "grey30")+
  theme(legend.position = "none",strip.text.x = element_text(size = 14, face = "bold"), 
        axis.text.x = element_text(face = "bold"), strip.background = element_rect(fill = "grey98"), 
        axis.title.x = element_text(face = "bold"))+
  scale_y_continuous(labels = scales::comma)+
  labs(x = "Title Sentiment", y = "", color = "")+
  facet_wrap(~name, scales = "free")
```

The way to interpret this plot is that the "violin" part shows the distribution of the data (bigger bulge where there are more data points) and the boxplot shows the median (bar in the middle) and first/third quartile (region containing the middle 75% of the data). We see that engagement (likes and comments) seem to be higher for videos with positive or negative title sentiments than neutral sentiments. We also see that views are on average higher for videos with negative titles. In order to take a closer look at this, I will next look at views in isolation, this time filtering out any outliers. 


```{r views-violin-plot}
ggthemr(palette = "pale", layout = "scientific")
#box and violin plot of just views with outliers removed
video_data %>%
  select(view_count, my_ratings) %>%
  rename(Views = view_count)%>%
  filter(Views <= 2.2e7) %>%
  ggplot(aes(x = my_ratings, y = Views))+
  geom_violin(aes(fill = my_ratings))+
  geom_boxplot(width = .2, fill = "grey95", color = "grey30")+
  scale_y_continuous(labels = scales::comma)+
  theme(legend.position = "none",  axis.text.x = element_text(face = "bold"), 
        axis.title = element_text(face = "bold"))+
  labs(x = "Title Sentiment", y = "Views", color = "")
```

These visualizations provide some support for the idea that more negative titles result in more video views on average. Both median and mean views are higher when the title is negative.


Next, it would be nice to examine if there has been a change in views, comments, or likes throughout the history of the channel. 

```{r timeseries, fig.width = 7, fig.height = 8}
ggthemr(palette = "fresh")
timeseries <- video_data %>%
  select(vid_title, view_count, like_count, comment_count, upload_date) %>%
  rename(Views = view_count, Likes = like_count, Comments = comment_count)%>%
  pivot_longer(cols = c("Views", "Likes", "Comments")) %>%
  ggplot( aes(x = upload_date, y = value))+
  geom_point(size = .5)+
  geom_line()+
  geom_smooth(method = "lm")+
  facet_wrap(~name, scale = "free", ncol = 1)+
  labs(x = "Upload Date", y = "")+
  scale_y_continuous(labels = scales::comma)+
  theme(strip.text.x = element_text(size = 12, face = "bold"), 
        strip.background = element_rect(fill = "grey97"), axis.text.x = element_text(face = "bold"), 
        axis.title.x = element_text(face = "bold"))

timeseries


#interactive version: uncomment and run if desired
#ggplotly(timeseries)
```

Views have been pretty constant through time, but likes and comments have increased notably over time. This might give some insight into how viewers are engaging with the videos - with new viewers liking and commenting on newer videos, and then more passively engaging with older content as they go back through the back catalog of videos.



```{r}
video_data %>%
  select(view_count, my_ratings) %>%
  group_by(my_ratings)%>%
  summarize(mean_views = mean(view_count))%>%
  formattable

video_data %>%
  select(comment_count, my_ratings) %>%
  group_by(my_ratings)%>%
  summarize(mean_comments = mean(comment_count))
```









Abandoned plots---------------------------------



```{r}
#column plot of mean likes, views, and comments for different title sentiments (model)
video_data %>%
  select(view_count, like_count, comment_count, title_sentiment) %>%
  group_by(title_sentiment)%>%
  summarize(Views = mean(view_count), Likes = mean(like_count), Comments = mean(comment_count)) %>%
  pivot_longer(cols = c("Views", "Likes", "Comments")) %>%
  ggplot(aes(x = title_sentiment, y = value, fill = title_sentiment))+
  geom_col()+
  theme_bw()+
  theme(legend.position = "none")+
  labs(x = "Title Sentiment", y = "", color = "")+
  facet_wrap(~name, scales = "free")

#column plot of mean likes, views, and comments for different title sentiments (human)
video_data %>%
  select(view_count, like_count, comment_count, my_ratings) %>%
  group_by(my_ratings)%>%
  summarize(Views = mean(view_count), Likes = mean(like_count), Comments = mean(comment_count)) %>%
  pivot_longer(cols = c("Views", "Likes", "Comments")) %>%
  ggplot(aes(x = my_ratings, y = value, fill = my_ratings))+
  geom_col()+
  theme_bw()+
  theme(legend.position = "none")+
  labs(x = "Title Sentiment", y = "", color = "")+
  facet_wrap(~name, scales = "free")
```


```{r}
#violin plot of model data
ggthemr(palette = "pale", layout = "scientific")
video_data %>%
  select(view_count, like_count, comment_count, title_sentiment) %>%
  rename(Views = view_count, Likes = like_count, Comments = comment_count)%>%
  pivot_longer(cols = c("Views", "Likes", "Comments")) %>%
  ggplot(aes(x = title_sentiment, y = value))+
  geom_violin(aes(fill = title_sentiment))+
  geom_boxplot(width = .2, fill = "grey95", color = "grey30")+
  theme(legend.position = "none")+
  labs(x = "Title Sentiment", y = "", color = "")+
  scale_y_continuous(labels = scales::comma)+
  facet_wrap(~name, scales = "free")
```


